---
title: "Reproduceable Research - Week 2 Assignment"
author: "Nils"
date: "25 September 2016"
output: html_document
---

```{r setup, include=FALSE, ref.label = TRUE,}
 opts_chunk$set(echo = TRUE)

```

# Submission Week 2 - Making research reproducable

## Commit containing full submission

1. Code for reading in the dataset and/or processing the data
2. Histogram of the total number of steps taken each day
3. Mean and median number of steps taken each day
4. Time series plot of the average number of steps taken
5. The 5-minute interval that, on average, contains the maximum number of steps
6. Code to describe and show a strategy for imputing missing data
7. Histogram of the total number of steps taken each day after missing values are imputed
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

These steps will now be performed one by one in the following subchapters.

## 1. Read from the file.

The working dorectory is already set. Therefore the file can be read with the following code:

```{r read data}
setwd("C:/Users/nilsn/Google Drive/Weiterbildung/R Programming/reproducablaresearch/week 2")
input <- read.csv("activity.csv")
input <- as.data.frame(input)
input$date <- as.Date(as.character(input$date), "%Y-%m-%d")
input$steps <- as.numeric(input$steps)
input$interval <- as.numeric(input$interval)
```

Additionally the table is formatted as a data frame and the date column is formated as date.

## 2. Histogram of the total number of steps taken each day

As the summary for the date shows, each date includes 288 measurements. Therefore the input has to be aggregated by date to sum up the total steps per date.
NA values are interpreted here as no steps.

```{r}
summary(input$date)
no1 <- aggregate(input$steps,by = list(date=input$date), FUN = sum, na.rm = TRUE)
hist(no1$x, xlab = "Steps per day", main = "Total number of steps taken each day")
```

## 3. Mean and median number of steps taken each day

The next figure shows the mean & median of steps each day, if NA values are interpreted as 0 steps.

```{r}
mean(no1$x)
median(no1$x)
```

## 4. Time series plot of the average number of steps taken

The first step is to aggreagte the data on mean per day, the next step is to plot the data.

```{r}
no4 <- aggregate(input$steps,by = list(date=input$date), FUN = mean, na.rm = TRUE)
plot(no4$date, no4$x, xlab = "date", ylab = "steps", main = "Average steps per day within an interval")
```



## 5. The 5-minute interval that, on average, contains the maximum number of steps


```{r}
no5 <- aggregate(input$steps,by = list(interval=input$interval), FUN = mean, na.rm = TRUE)
plot(no5$interval, no5$x, xlab = "interval", ylab = "steps", main = "Average steps per interval")
library(dplyr)
filter(no5, x == max(x))
```

As the graph shows, the maximum lies by 'r filter(no5, x == max(x))'.

## 6. Code to describe and show a strategy for imputing missing data

Instead of using the claasical imputation code of impute, this data set uses the mice package to impute the missing values.

For this reason in a first step, all variables need to be converted to numeric vectors. Afterwards the relevant items are imputed.

```{r, cache = TRUE, results = 'hide'}
input2 <- input
input2$date <- as.numeric(input2$date)
library(mice)
full <- mice(input2, m=3, maxit = 50, meth = "pmm", seed = 500)
completedData <- complete(full,1)
```

## 7. Histogram of the total number of steps taken each day after missing values are imputed

The completed data set is added as new columns to the input data. From this the histogram is done, similiar to task 2.

```{r}
summary(completedData$date)
input$steps2 <- completedData$steps
rm(input2, completedData)
no7 <- aggregate(input$steps2,by = list(date=input$date), FUN = sum, na.rm = TRUE)
hist(no7$x, xlab = "Steps per day", main = "Total number of steps taken each day")
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


```{r}
input$weekday <- weekdays(input$date)
weedays <- unique(input$weekday)[1:5]
weekends <- unique(input$weekday)[6:7]
aveweekend <- input[which(input$weekday=="Samstag" |input$weekday=="Sonntag"),]
avewd <- mean(aveweekend$steps2)
aveweekday <- input[which(!input$weekday=="Samstag" & !input$weekday=="Sonntag"),]
avewy <- mean(aveweekday$steps2)
par(mfrow = c(1, 2))
boxplot(aveweekday$steps2, ylim = c(0,40))
boxplot(aveweekend$steps2, ylim = c(0,40))
```

## 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

```{r, eval = FALSE, results = 'hide'}
input <- as.data.frame(input)
input$date <- as.Date(as.character(input$date), "%Y-%m-%d")
input$steps <- as.numeric(input$steps)
input$interval <- as.numeric(input$interval)
summary(input$date)
no1 <- aggregate(input$steps,by = list(date=input$date), FUN = sum, na.rm = TRUE)
hist(no1$x, xlab = "Steps per day", main = "Total number of steps taken each day")
mean(no1$x)
median(no1$x)
no4 <- aggregate(input$steps,by = list(date=input$date), FUN = mean, na.rm = TRUE)
plot(no4$date, no4$x, xlab = "date", ylab = "steps", main = "Average steps per day within an interval")
no5 <- aggregate(input$steps,by = list(interval=input$interval), FUN = mean, na.rm = TRUE)
plot(no5$interval, no5$x, xlab = "interval", ylab = "steps", main = "Average steps per interval")
library(dplyr)
filter(no5, x == max(x))
input2 <- input
input2$date <- as.numeric(input2$date)
library(mice)
full <- mice(input2, m=3, maxit = 50, meth = "pmm", seed = 500)
completedData <- complete(full,1)
summary(completedData$date)
input$steps2 <- completedData$steps
rm(input2, completedData)
no7 <- aggregate(input$steps2,by = list(date=input$date), FUN = sum, na.rm = TRUE)
hist(no7$x, xlab = "Steps per day", main = "Total number of steps taken each day")
input$weekday <- weekdays(input$date)
weedays <- unique(input$weekday)[1:5]
weekends <- unique(input$weekday)[6:7]
aveweekend <- input[which(input$weekday=="Samstag" |input$weekday=="Sonntag"),]
avewd <- mean(aveweekend$steps2)
aveweekday <- input[which(!input$weekday=="Samstag" & !input$weekday=="Sonntag"),]
avewy <- mean(aveweekday$steps2)
par(mfrow = c(1, 2))
boxplot(aveweekday$steps2, ylim = c(0,40))
boxplot(aveweekend$steps2, ylim = c(0,40))
```

















